\chapter{Appendix B: Interview Guide}
\label{cha:anhang_B}

\pagenumbering{arabic} % Reset Page-Numbering to 1
\renewcommand*{\thepage}{B-\arabic{page}} % Nummerierung B-1

The following interview guide was used to interview the experts, as described in \Cref{sec:expert_evaluation_methodology}.

\subsection*{Introduction and Purpose}
\begin{itemize}
    \item \textbf{Brief Introduction:} Provide a concise overview of the research context through a short slide presentation, ensuring the expert fully understands the study’s objectives and scope before proceeding with the interview.
    \item \textbf{Objective:} This interview aims to gather insights on the implementation of \acrlong{dpp} systems, including industry requirements, technology adoption, and preferences for foundational technologies and data interfaces.
    \item \textbf{Scope:} The findings will help evaluate and score various technologies against \acrlong{kpi}s (\ac{kpi}s) and may guide the design of a pilot project.
    \item \textbf{Confidentiality:} Privacy or anonymization requests will be respected. With the expert's permission, the interview will be recorded and transcribed for research purposes.
\end{itemize}

\subsection*{Part 1: Background Information}
\begin{enumerate}
    \item \textbf{About the Expert}
    \begin{enumerate}
        \item \textbf{Professional Background:} 
        \begin{itemize}
            \item ``What is your professional background and role in your organization?''
        \end{itemize}
        \item \textbf{Experience with \ac{dpp} Technologies:}
        \begin{itemize}
            \item ``What is your experience with implementing or managing technologies related to \ac{dpp}?''
        \end{itemize}
    \end{enumerate}

    \item \textbf{About the Organization}
    \begin{enumerate}
        \item \textbf{Role in the \ac{dpp} Ecosystem:}
        \begin{itemize}
            \item ``What is your organization’s role in the \ac{dpp} ecosystem?''
        \end{itemize}
    \end{enumerate}
\end{enumerate}

\subsection*{Part 2: Requirements and Industry Needs}

\begin{enumerate}
    \item \textbf{Core Requirements for \ac{dpp} Systems}
    \begin{enumerate}
        \item \textbf{Overview of \ac{kpi}s:} Present and briefly define each of the following:
        \begin{itemize}
            \item Interoperability
            \item Scalability
            \item Data Security and Integrity
            \item Portability
            \item Modularity
            \item Usability
            \item Reliability
            \item Cost-Effectiveness
            \item Compliance
        \end{itemize}
        \item \textbf{Initial Feedback:} 
        \begin{itemize}
            \item ``Do you agree with these \ac{kpi}s as critical for \ac{dpp} systems?''
        \end{itemize}
        \item \textbf{Additional \ac{kpi}s:} 
        \begin{itemize}
            \item ``Are there any additional \ac{kpi}s you believe are important but not included here? Could you provide examples?''
        \end{itemize}
        \item \textbf{Relevance:} 
        \begin{itemize}
            \item ``How does your organization weigh these factors when evaluating or implementing \ac{dpp}-related technologies?''
        \end{itemize}
        \item \textbf{Ranking Exercise:} 
        \begin{itemize}
            \item ``Let’s rank these \ac{kpi}s according to importance.'' (Use a predefined table to capture responses.)
        \end{itemize}
        \item \textbf{Contextual Differences:} 
        \begin{itemize}
            \item ``Do these requirements differ significantly across industries or product categories? Could you share specific examples?''
        \end{itemize}
    \end{enumerate}

    \item \textbf{Regulatory and Standardization Impact}
    \begin{enumerate}
        \item \textbf{Influence of \ac{espr}:}
        \begin{itemize}
            \item ``How do regulatory frameworks like \ac{espr} influence your technology decisions?''
        \end{itemize}
        \item \textbf{Key \ac{espr} Requirements:} 
        \begin{itemize}
            \item Traceability and transparency of product lifecycle data.
            \item Durability and repairability to reduce waste and extend product life.
            \item Recyclability and efficient end-of-life management.
            \item Accessibility of data for all stakeholders.
            \item Compliance through mandatory \ac{dpp}s for regulated product categories.
        \end{itemize}
        \item \textbf{Standardized vs. Individualized Solutions:}
        \begin{itemize}
            \item ``Do you think it’s more important to adopt standardized approaches or use individualized solutions for \ac{dpp} implementation? Why?''
        \end{itemize}
    \end{enumerate}
\end{enumerate}

\subsection*{Part 3: Technology Evaluation and Preferences}

\begin{enumerate}
    \item \textbf{Technology Affinities}
    \begin{enumerate}
        \item \textbf{Current Use:}
        \begin{itemize}
            \item ``Which technologies (e.g., blockchain, cloud, federated systems) are currently being used or considered for \ac{dpp}s in your organization? Why were these chosen?''
        \end{itemize}
        \item \textbf{Driving Factors:}
        \begin{itemize}
            \item ``What factors drive the selection of these technologies (e.g., cost, ease of integration, scalability)? Which \ac{kpi}s are most influential here?''
        \end{itemize}
    \end{enumerate}
    
    \item \textbf{Challenges with Current Technologies}
    \begin{enumerate}
        \item \textbf{Major Pain Points:}
        \begin{itemize}
            \item ``What are the most significant challenges you face with the current technology stack (e.g., high costs, lack of standards, technical complexity)?''
        \end{itemize}
        \item \textbf{Mitigation Strategies:}
        \begin{itemize}
            \item ``How are these challenges being addressed or mitigated within your organization?''
        \end{itemize}
    \end{enumerate}
\end{enumerate}

\subsection*{Part 4: Future Potential and Trends}

\begin{enumerate}
    \item \textbf{Scaling Potentials}
    \begin{enumerate}
        \item \textbf{5--10 Year Outlook:}
        \begin{itemize}
            \item ``How do you envision the scalability of \ac{dpp} systems across industries over the next 5--10 years?''
        \end{itemize}
        \item \textbf{Emerging Innovations:}
        \begin{itemize}
            \item ``Are there emerging technologies or trends that could revolutionize \ac{dpp} implementation? Have you encountered any pilot projects or frameworks worth noting?''
        \end{itemize}
    \end{enumerate}

    \item \textbf{Industry Affinities}
    \begin{itemize}
        \item ``Which technologies or approaches do you think are becoming standard across the \ac{dpp} ecosystem? Why?''
    \end{itemize}

    \item \textbf{Best Practices and Recommendations}
    \begin{enumerate}
        \item \textbf{Guidelines:}
        \begin{itemize}
            \item ``Based on your expertise, what best practices should organizations follow for selecting and implementing \ac{dpp} technologies?''
        \end{itemize}
        \item \textbf{Common Pitfalls:}
        \begin{itemize}
            \item ``What common pitfalls have you observed in \ac{dpp} implementations, and how can they be avoided?''
        \end{itemize}
    \end{enumerate}
\end{enumerate}

\subsection*{Closing}
\begin{itemize}
    \item \textbf{Follow-up:} Request permission to reach out for any clarifications.
    \item \textbf{Appreciation:} Thank the interviewee for their time, insights, and support in contributing to this research.
\end{itemize}